#Course 5 Reproducible Research - Peer grade assignment 1
## Activity Monitoring device analysis
###Date: 11/01/2017
###Coded by: Vidhya Swaminathan

### Read dataset into dataframe for processing

```{r Create Dataframe, echo=TRUE}
setwd("C:/Temp/Coursera/ReproducibleResearch/Wk2Prj")

if(!file.exists("activity.csv")) { unzip("repdata_data_activity.zip") }

activity_ds = read.csv("activity.csv")

```

### Create a histogram of the total number of steps taken each day
```{r Create Histogram/Date, echo=TRUE}
activity_steps_by_day <- aggregate(steps ~ date, activity_ds, sum, na.action = na.pass)
hist(activity_steps_by_day$steps, col="orange", main = "Total no steps/day", xlab = "No of steps")

```

### Mean and median number of steps taken each day
```{r Mean/ Median on steps each day, echo=TRUE}
r_steps_mean <- mean(activity_steps_by_day$steps)
message("Mean = ", r_steps_mean)

r_steps_medn <- median(activity_steps_by_day$steps)
message("Median = ", r_steps_medn)
```

### Create a Time series plot for the average number of steps taken. 

```{r time series, echo=TRUE}
activity_mean_interval <- tapply(activity_ds$steps, activity_ds$interval, mean, na.rm = T)

plot(activity_mean_interval, type = "l", main = "Time Series Plot", xlab = "The 5 minute interval", ylab = "The Average no of steps")

```

### Pick the first 10 data for the 5-minute interval that, on average, contains the maximum number of steps

```{r max no. of steps, echo=TRUE}
head(sort(activity_mean_interval, decreasing = TRUE), 1)
```

### To compute and show a strategy for imputing missing data

```{r, echo=TRUE}

activity_stepsNA <- sum(is.na(activity_ds$steps))
activity_dateNA <- sum(is.na(activity_ds$date))
activity_intvlNA <- sum(is.na(activity_ds$interval))

message("Missing Steps = ", activity_stepsNA)
message("Missing Dates = ", activity_dateNA)
message("Missing Interval = ", activity_intvlNA)

table(is.na(activity_ds$steps), activity_ds$interval, activity_ds$date)

```

### Histogram graph for the total number of steps taken each day after imputing missing data
```{r, echo=TRUE}
activity_steps_by_day <- aggregate(steps ~ date, activity_ds, sum, na.action = na.exclude)
hist(activity_steps_by_day$steps, col="orange", main = "Total no steps/day", xlab = "No. of steps")
```

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r comparison, echo=TRUE}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

activity_ds$dow = as.factor(ifelse(is.element(weekdays(as.Date(activity_ds$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_activity <- aggregate(steps ~ interval + dow, activity_ds, mean)

library(lattice)

xyplot(steps_by_interval_activity$steps ~ steps_by_interval_activity$interval|steps_by_interval_activity$dow, main="Average steps/day by interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```

### End of document
